---
title: "Project_stat"
author: "I.Caria, C.De Luca, M.Pernini"
date: "13/06/2022"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

## Libraries

```{r message=FALSE, warning=FALSE}

library(ggplot2)
library(tidyverse)
library(leaps)
library(ggcorrplot)
library(regclass)
library(boot)
library(caret)
library(MASS)
library(knitr)
library(corrplot)
library(glmnet)
library(plotly)

```


## Hotel booking demand dataset

We decided to analyze the *Hotel booking demand dataset* that we load from Kaggle. This dataset contains information about two different kinds of hotel: City Hotel and Resort Hotel. Each observation represents an ho.tel booking. Both hotels are located in Portugal: the resort hotel at the resort region of Algarve and the city hotel at the city of Lisbon.


```{r}
# Load the dataset

hotel_bookings <- read.csv("hotel_bookings.csv", na.strings="NULL")
View(hotel_bookings)
```


# Dataset Pre-Processing

The dataset contains 32 variables describing 119390 observations.

```{r}
# First look to the dataset

glimpse(hotel_bookings)

```

As we can see from the code above, there are many character variables that we converted into factors. Furthermore, we noticed that some categorical variables like *children* were numeric, so we converted them.

```{r warning=FALSE}
# Convert character columns into factors

hotel_bookings_new <- as.data.frame(unclass(hotel_bookings),
                       stringsAsFactors = TRUE)

# Convert binary columns "is_canceled" and "is_repeated_guest" into factor

hotel_bookings_new$is_canceled <- as.factor(hotel_bookings_new$is_canceled)
levels(hotel_bookings_new$is_canceled) <- c(0, 1)

hotel_bookings_new$is_repeated_guest <- as.factor(hotel_bookings_new$is_repeated_guest)
levels(hotel_bookings_new$is_repeated_guest) <- c(0, 1)

# Convert column "arrival_date_year" into factor

hotel_bookings_new$arrival_date_year <- as.factor(hotel_bookings_new$arrival_date_year)
levels(hotel_bookings_new$arrival_date_year) <- c("2015", "2016", "2017")

# Convert column "children" into numeric

hotel_bookings_new$children <- as.numeric(as.character(hotel_bookings_new$children))

# Convert column "reservation status date" into date

hotel_bookings_new$reservation_status_date <- as.Date(
  hotel_bookings_new$reservation_status_date, format = "%Y-%m-%d")
```

The dataset provides two different variables for the stay: *stays_in_weekend_nights* and *stays_in_week_nights*. We decided to add the sum of these two variables as a new variable *total_stays* for ease of analyses.

```{r}
# New column for total stays

hotel_bookings_new=hotel_bookings_new%>%
  mutate(total_stays=(stays_in_week_nights + stays_in_weekend_nights) )
```

### Missing values

```{r}

colSums(is.na(hotel_bookings_new))[colSums(is.na(hotel_bookings_new))>0]
```

Since there are only 4 Nan values for the variable *children*, we decided to replace them with the value 0.
The variables *agent* and *company* have too many Nan values, therefore we removed them. We left untouched the variable *country* because we did not use it in our models.

```{r}
# Replacing missing values in children column from the corresponding babies column

n_children <- length(hotel_bookings_new$children)
for (i in 1:n_children) {
  if (is.na(hotel_bookings_new$children[i]))
    hotel_bookings_new$children[i] <- 0
}

# Remove columns "agent" and "company"

index_agent <- which(colnames(hotel_bookings_new)=="agent")
index_company <- which(colnames(hotel_bookings_new)=="company")
hotel_bookings_new = hotel_bookings_new[-c(index_agent, index_company)]

```

At the end of the pre-processing, we obtained the following dataset:

```{r echo=FALSE}
# Summary

summary(hotel_bookings_new)
```

# EDA

As stated above,the dataset contains information about two different kinds of hotel: City Hotel and Resort Hotel. There are 79330 observation for the former and 40060 for the latter.


```{r}
# Hotel donut plot

df_hotel <- as.data.frame(hotel_bookings_new[, c("hotel")])
df_hotel <- as.data.frame(lapply(df_hotel, function(x) as.data.frame(table(x))))

colnames(df_hotel) <- c("hotel", "frequency")

df_hotel

colors_donut <- c('rgb((119,51,68))','rgb((227,181,164))') 

fig_hotel <- df_hotel %>% plot_ly(labels = ~hotel, values = ~frequency, 
                                  marker = list(colors = colors_donut,
                                  line = list(color = '#FFFFFF', width = 1)))
fig_hotel <- fig_hotel %>% add_pie(hole = 0.6)
fig_hotel <- fig_hotel %>% layout(title = "Total number of booking for each hotel",  
                      showlegend = T,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig_hotel

```


Both hotels are located in Portugal; this is the reason why most of the guests come from Portugal, as we can see from the map plot below:


```{r}
# Country plot

df_country <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0, c("country")])
df_country <- as.data.frame(lapply(df_country, function(x) as.data.frame(table(x))))
colnames(df_country) <- c('country', 'frequency')

# we used log-scale to see better the different number of guests for each country
fig_country <- plot_ly(df_country, type='choropleth', locations=df_country$country, 
                       z=log(df_country$frequency+1), colorscale = 'Viridis' )
fig_country <- fig_country %>% layout(title = "Country plot")
fig_country
```

Total bookings considering both canceled and not canceled
Total guests considering only is_canceled==0


```{r, fig.dim=c(10,8)}
# Total bookings for each hotel by month of arrival date

df_months_City <- as.data.frame(hotel_bookings_new[hotel_bookings_new$hotel=='City Hotel', 
                                                   c( "arrival_date_month")])

df_months_Resort <- as.data.frame(hotel_bookings_new[hotel_bookings_new$hotel=='Resort Hotel', 
                                                     c( "arrival_date_month")])

df_months_City <- as.data.frame(lapply(df_months_City, function(x) as.data.frame(table(x))))

df_months_Resort <- as.data.frame(lapply(df_months_Resort, function(x) as.data.frame(table(x))))

colnames(df_months_City) <- c("arrival_date_month_City", "frequency_City" )

colnames(df_months_Resort) <- c("arrival_date_month_Resort", "frequency_Resort")

df_months_City$arrival_date_month_City <- factor(df_months_City$arrival_date_month_City, 
          levels = c("January", "February", "March","April", "May", "June", "July", 
                     "August", "September", "October", "November", "December"))

df_months_Resort$arrival_date_month_Resort <- factor(df_months_Resort$arrival_date_month_Resort, 
          levels = c("January", "February", "March","April", "May", "June", "July", 
                     "August", "September", "October", "November", "December"))


fig_months <- plot_ly()

fig_months <- fig_months %>% add_lines(data=df_months_City, 
                                       x = ~arrival_date_month_City, y = ~frequency_City, 
                                       name = 'City Hotel', type = 'scatter', mode = 'lines', 
                                       line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_months <- fig_months %>% add_lines(data=df_months_Resort, x = ~arrival_date_month_Resort, 
                                       y = ~frequency_Resort, name = 'Resort Hotel', 
                                       type = 'scatter', mode = 'lines', 
                                       line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_months <- fig_months %>% 
layout(title = "Total bookings for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total bookings"))

fig_months
```

For each year 


```{r, fig.dim=c(10,8)}
# Total bookings for each hotel by month (year: 2015)

df_months_City_2015 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$hotel=='City Hotel' 
              & hotel_bookings_new$arrival_date_year==2015, c( "arrival_date_month")])

df_months_Resort_2015 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$hotel=='Resort Hotel'
              & hotel_bookings_new$arrival_date_year==2015, c( "arrival_date_month")])

df_months_City_2015 <- as.data.frame(lapply(df_months_City_2015, function(x) 
  as.data.frame(table(x))))

df_months_Resort_2015 <- as.data.frame(lapply(df_months_Resort_2015, function(x) 
  as.data.frame(table(x))))

colnames(df_months_City_2015) <- c("arrival_date_month_City", "frequency_City" )

colnames(df_months_Resort_2015) <- c("arrival_date_month_Resort", "frequency_Resort")

df_months_City_2015$arrival_date_month_City <- factor(df_months_City_2015$arrival_date_month_City, 
            levels = c("January", "February", "March","April", "May", "June", "July", 
                       "August", "September", "October", "November", "December"))

df_months_Resort_2015$arrival_date_month_Resort <- factor(df_months_Resort_2015$arrival_date_month_Resort, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_months_2015 <- plot_ly()

fig_months_2015 <- fig_months_2015 %>% add_lines(data=df_months_City_2015, x = ~arrival_date_month_City, y = ~frequency_City, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_months_2015 <- fig_months_2015 %>% add_lines(data=df_months_Resort_2015, x = ~arrival_date_month_Resort, y = ~frequency_Resort, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_months_2015 <- fig_months_2015 %>% layout(title = "2015 - Total bookings for each hotel by month of arrival date",
         xaxis = list(title = "Months",tickangle = -45),
         yaxis = list (title = "Total bookings"))
fig_months_2015

```

```{r, fig.dim=c(10,8)}

# Total bookings for each hotel by month (year: 2016)

df_months_City_2016 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2016, c( "arrival_date_month")])

df_months_Resort_2016 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$hotel=='Resort Hotel'& hotel_bookings_new$arrival_date_year==2016, c( "arrival_date_month")])

df_months_City_2016 <- as.data.frame(lapply(df_months_City_2016, function(x) as.data.frame(table(x))))

df_months_Resort_2016 <- as.data.frame(lapply(df_months_Resort_2016, function(x) as.data.frame(table(x))))

colnames(df_months_City_2016) <- c("arrival_date_month_City", "frequency_City" )

colnames(df_months_Resort_2016) <- c("arrival_date_month_Resort", "frequency_Resort")

df_months_City_2016$arrival_date_month_City <- factor(df_months_City_2016$arrival_date_month_City, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))

df_months_Resort_2016$arrival_date_month_Resort <- factor(df_months_Resort_2016$arrival_date_month_Resort, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_months_2016 <- plot_ly()

fig_months_2016 <- fig_months_2016 %>% add_lines(data=df_months_City_2016, x = ~arrival_date_month_City, y = ~frequency_City, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_months_2016 <- fig_months_2016 %>% add_lines(data=df_months_Resort_2016, x = ~arrival_date_month_Resort, y = ~frequency_Resort, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_months_2016 <- fig_months_2016 %>% layout(title = "2016 - Total bookings for each hotel by month of arrival date",
         xaxis = list(title = "Months",tickangle = -45),
         yaxis = list (title = "Total bookings"))

fig_months_2016
```

```{r, fig.dim=c(10,8)}


# Total bookings for each hotel by month (year: 2017)

df_months_City_2017 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2017, c( "arrival_date_month")])

df_months_Resort_2017 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$hotel=='Resort Hotel'& hotel_bookings_new$arrival_date_year==2017, c( "arrival_date_month")])

df_months_City_2017 <- as.data.frame(lapply(df_months_City_2017, function(x) as.data.frame(table(x))))

df_months_Resort_2017 <- as.data.frame(lapply(df_months_Resort_2017, function(x) as.data.frame(table(x))))

colnames(df_months_City_2017) <- c("arrival_date_month_City", "frequency_City" )

colnames(df_months_Resort_2017) <- c("arrival_date_month_Resort", "frequency_Resort")

df_months_City_2017$arrival_date_month_City <- factor(df_months_City_2017$arrival_date_month_City, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))

df_months_Resort_2017$arrival_date_month_Resort <- factor(df_months_Resort_2017$arrival_date_month_Resort, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_months_2017 <- plot_ly()

fig_months_2017 <- fig_months_2017 %>% add_lines(data=df_months_City_2017, x = ~arrival_date_month_City, y = ~frequency_City, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_months_2017 <- fig_months_2017 %>% add_lines(data=df_months_Resort_2017, x = ~arrival_date_month_Resort, y = ~frequency_Resort, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_months_2017 <- fig_months_2017 %>% layout(title = "2017 - Total bookings for each hotel by month of arrival date",
         xaxis = list(title = "Months",tickangle = -45),
         yaxis = list (title = "Total bookings"))

fig_months_2017
```








Total guests








```{r}
# Total guests for each hotel by month

df_City <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel', c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_City <- df_City %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 

df_Resort <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel', c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_Resort <- df_Resort %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 


df_guests_City$arrival_date_month <- factor(df_guests_City$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))

df_guests_Resort$arrival_date_month <- factor(df_guests_Resort$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_guests <- plot_ly()

fig_guests <- fig_guests %>% add_lines(data=df_guests_City, x = ~arrival_date_month, y = ~guests, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_guests <- fig_guests %>% add_lines(data=df_guests_Resort, x = ~arrival_date_month, y = ~guests, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_guests <- fig_guests %>% layout(title = "Total guests for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total guests"))

fig_guests

```



```{r}
# Total guests for each hotel by month (year: 2015)

df_2015_City <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2015, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2015_City <- df_2015_City %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 

df_2015_Resort <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel' & hotel_bookings_new$arrival_date_year==2015, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2015_Resort <- df_2015_Resort %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 


df_guests_2015_City$arrival_date_month <- factor(df_guests_2015_City$arrival_date_month, levels = c("July", "August", "September", "October", "November", "December"))

df_guests_2015_Resort$arrival_date_month <- factor(df_guests_2015_Resort$arrival_date_month, levels = c("July", "August", "September", "October", "November", "December"))


fig_guests_2015 <- plot_ly()

fig_guests_2015 <- fig_guests_2015 %>% add_lines(data=df_guests_2015_City, x = ~arrival_date_month, y = ~guests, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_guests_2015 <- fig_guests_2015 %>% add_lines(data=df_guests_2015_Resort, x = ~arrival_date_month, y = ~guests, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_guests_2015 <- fig_guests_2015 %>% layout(title = "2015 - Total guests for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total guests"))

fig_guests_2015

```




```{r}
# Total guests for each hotel by month (year: 2016)

df_2016_City <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2016, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2016_City <- df_2016_City %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 

df_2016_Resort <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel' & hotel_bookings_new$arrival_date_year==2016, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2016_Resort <- df_2016_Resort %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 


df_guests_2016_City$arrival_date_month <- factor(df_guests_2016_City$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))

df_guests_2016_Resort$arrival_date_month <- factor(df_guests_2016_Resort$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_guests_2016 <- plot_ly()

fig_guests_2016 <- fig_guests_2016 %>% add_lines(data=df_guests_2016_City, x = ~arrival_date_month, y = ~guests, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_guests_2016 <- fig_guests_2016 %>% add_lines(data=df_guests_2016_Resort, x = ~arrival_date_month, y = ~guests, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_guests_2016 <- fig_guests_2016 %>% layout(title = "2016 - Total guests for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total guests"))

fig_guests_2016
```



```{r}
# Total guests for each hotel by month (year: 2017)

df_2017_City <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2017, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2017_City <- df_2017_City %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 

df_2017_Resort <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel' & hotel_bookings_new$arrival_date_year==2017, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2017_Resort <- df_2017_Resort %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 


df_guests_2017_City$arrival_date_month <- factor(df_guests_2017_City$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August"))

df_guests_2017_Resort$arrival_date_month <- factor(df_guests_2017_Resort$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August"))


fig_guests_2017 <- plot_ly()

fig_guests_2017 <- fig_guests_2017 %>% add_lines(data=df_guests_2017_City, x = ~arrival_date_month, y = ~guests, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_guests_2017 <- fig_guests_2017 %>% add_lines(data=df_guests_2017_Resort, x = ~arrival_date_month, y = ~guests, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_guests_2017 <- fig_guests_2017 %>% layout(title = "2016 - Total guests for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total guests"))

fig_guests_2017
```


Reserved room type vs Assigned room type

```{r}
# Reserved room type table

table_hotel_reserved <- table(hotel_bookings_new$hotel, 
                              hotel_bookings_new$reserved_room_type)
table_hotel_reserved
```

```{r}
# Reserved room type plot

city_reserved <- table_hotel_reserved[c('City Hotel'),]
resort_reserved <- table_hotel_reserved[c('Resort Hotel'),]
room_type_reserved <- colnames(table_hotel_reserved)

df_reserved <- data.frame(room_type_reserved, city_reserved, resort_reserved)

fig_reserved <- plot_ly(df_reserved, x = ~room_type_reserved, y = ~city_reserved, 
                        type = 'bar', name = 'City Hotel', 
                        marker = list(color = 'rgb(239,115,101)'))
fig_reserved <- fig_reserved %>% add_trace(y = ~resort_reserved, name = 'Resort Hotel', 
                                           marker = list(color='rgb(251,210,108)'))
fig_reserved <- fig_reserved %>% layout(title = "Reserved room type", 
                                        xaxis = list(title = 'room type'), 
                                        yaxis =list(title = 'number of bookings'), 
                                        barmode = 'group')
fig_reserved
```

```{r}
# Assigned room type table

table_hotel_assigned <- table(hotel_bookings_new$hotel, 
                              hotel_bookings_new$assigned_room_type)
table_hotel_assigned
```

```{r}
# Assigned room type plot

city_assigned <- table_hotel_assigned[c('City Hotel'),]
resort_assigned <- table_hotel_assigned[c('Resort Hotel'),]
room_type_assigned <- colnames(table_hotel_assigned)

df_assigned <- data.frame(room_type_assigned, city_assigned, resort_assigned)

fig_assigned <- plot_ly(df_assigned, x = ~room_type_assigned, y = ~city_assigned, 
                        type = 'bar', name = 'City Hotel', 
                        marker = list(color = 'rgb(239,115,101)'))
fig_assigned <- fig_assigned %>% add_trace(y = ~resort_assigned, name = 'Resort Hotel', 
                                           marker = list(color='rgb(251,210,108)'))
fig_assigned <- fig_assigned %>% layout(title = "Assigned room type", 
                                        xaxis = list(title = 'room type'), 
                                        yaxis =list(title = 'number of bookings'), 
                                        barmode = 'group')
fig_assigned
```



## Models

```{r, warning=FALSE}
complete_model<-glm(is_canceled~hotel+lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests,  family="binomial")
summary(complete_model)
model_AIC<-stepAIC(complete_model, direction = "both")
```

### AUC Curve
```{r}
logistic.prob_AIC=predict(model_AIC, type="response")
roc.out_AIC <- roc(hotel_bookings_new$is_canceled, logistic.prob_AIC, levels=c(0, 1)) 
plot(roc.out_AIC,  print.auc=TRUE, legacy.axes=TRUE, xlab="False positive rate", ylab="True positive rate", col="red") 
  
```
### Logistic VS LDA VS QDA
```{r}
#Train & Test
set.seed(321)
trainIndex <- createDataPartition(little_data$is_canceled,p=0.75,list=FALSE)
hotel_data_train=little_data[trainIndex,]
hotel_data_test=little_data[-trainIndex,]
is_canc_test=hotel_data_test$is_canceled
```


```{r}
#Logistic Regression
glm.fits <- glm(is_canceled~lead_time+adr+total_of_special_requests,data=hotel_data_train,family=binomial)
glm.probs <- predict(glm.fits,hotel_data_test,type="response")
glm.pred <- rep(0,29847)
glm.pred[glm.probs>.5] <- 1

#Evaluation
confusionMatrix(as.factor(glm.pred),is_canc_test)
cat("Mean error is:", mean(glm.pred!=is_canc_test))
```

```{r}
#Polinomial Regression
glm.fits <- glm(is_canceled~lead_time+adr+total_of_special_requests+I(adr^3)+I(lead_time^2)+I(total_of_special_requests^2),data=hotel_data_train,family=binomial)
glm.probs <- predict(glm.fits,hotel_data_test,type="response")
glm.pred <- rep(0,29847)
glm.pred[glm.probs>.5] <- 1

#Evaluation
confusionMatrix(as.factor(glm.pred),is_canc_test)
cat("Mean error is:", mean(glm.pred!=is_canc_test))
```

```{r}
#LDA
lda.fit <- lda(is_canceled~lead_time+adr+total_of_special_requests,data=hotel_data_train)
par(mar=c(1,1,1,1))
plot(lda.fit)
lda.pred <- predict(lda.fit, hotel_data_test)
lda.class <- lda.pred$class

#Evaluation
confusionMatrix(as.factor(lda.class),is_canc_test)
cat("Mean error is:", mean(glm.pred!=is_canc_test))
```

```{r}
#QDA
qda.fit <- qda(is_canceled~lead_time+adr+total_of_special_requests,data=hotel_data_train)
qda.pred <- predict(qda.fit, hotel_data_test)
qda.class <- qda.pred$class


# Evaluation
confusionMatrix(as.factor(qda.class),is_canc_test)
cat("Mean error is:", mean(glm.pred!=is_canc_test))
```

### LASSO

```{r}
set.seed(123)
training.samples <- hotel_bookings_new$is_canceled %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- hotel_bookings_new[training.samples, ]
test.data <- hotel_bookings_new[-training.samples, ]

## Dumy code categorical predictor variables
x <- model.matrix(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests, train.data)[,-1]
## Convert the outcome (class) to a numerical variable
y <- train.data$is_canceled
lambdas <- 10^seq(2, -3, by = -.1)
glmnet(x, y, family = "binomial", alpha = 0, lambda = NULL)

## Find the best lambda using cross-validation
set.seed(123) 
cv.lasso <- cv.glmnet(x, y, alpha = 1, family = "binomial")
## Fit the final model on the training data
model_lasso <- glmnet(x, y, alpha = 1, family = "binomial",
                lambda = cv.lasso$lambda.min)
## Display regression coefficients
coef(model_lasso)
## Make predictions on the test data
x.test <- model.matrix(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests, test.data)[,-1]
probabilities <- complete_model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
## Model accuracy
observed.classes <- test.data$is_canceled
#mean(predicted.classes == observed.classes)

cv.lasso <- cv.glmnet(x, y, alpha = 1, family = "binomial")
plot(cv.lasso)
cv.lasso$lambda.min
cv.lasso$lambda.1se
coef(cv.lasso, cv.lasso$lambda.min)
coef(cv.lasso, cv.lasso$lambda.1se)

## Final model with lambda.min
lasso.model <- glmnet(x, y, alpha = 1, family = "binomial",
                      lambda = cv.lasso$lambda.min)
## Make prediction on test data
x.test <- model.matrix(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests, test.data)[,-1]
probabilities <- lasso.model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
## Model accuracy
observed.classes <- test.data$is_canceled
#mean(predicted.classes == observed.classes)

confusionMatrix(as.factor(predicted.classes), observed.classes)


```

### RIDGE

```{r}
lambdas <- 10^seq(2, -3, by = -.1)
glmnet(x, y, family = "binomial", alpha = 0, lambda = NULL)

## Find the best lambda using cross-validation
set.seed(123) 
cv.ridge <- cv.glmnet(x, y, alpha = 0, family = "binomial")
## Fit the final model on the training data
model_ridge <- glmnet(x, y, alpha = 0, family = "binomial",
                lambda = cv.ridge$lambda.min)
## Display regression coefficients
coef(model_ridge)
## Make predictions on the test data
x.test <- model.matrix(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests, test.data)[,-1]
probabilities <- complete_model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
## Model accuracy
observed.classes <- test.data$is_canceled
#mean(predicted.classes == observed.classes)

#cv.ridge <- cv.glmnet(x, y, alpha = 1, family = "binomial")
plot(cv.ridge)
cv.ridge$lambda.min
cv.ridge$lambda.1se
coef(cv.ridge, cv.ridge$lambda.min)
coef(cv.ridge, cv.ridge$lambda.1se)

## Final model with lambda.min
ridge.model <- glmnet(x, y, alpha = 0, family = "binomial",
                      lambda = cv.ridge$lambda.min)
## Make prediction on test data
x.test <- model.matrix(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests, test.data)[,-1]
probabilities <- ridge.model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
## Model accuracy
observed.classes <- test.data$is_canceled
#mean(predicted.classes == observed.classes)

confusionMatrix(as.factor(predicted.classes), observed.classes)

```

