---
title: "Stat_Project"
author: "I.Caria, C.De Luca, M.Pernini"
date: "5/24/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---



```{r message=FALSE, warning=FALSE}
# Libraries

library(ggplot2)
library(tidyverse)
library(plotly)
```


We decided to analyze the *Hotel booking demand dataset* that we load from Kaggle. This dataset contains information about two different kinds of hotel: City Hotel and Resort Hotel. Each observation represents an hotel booking. Both hotels are located in Portugal: the resort hotel at the resort region of Algarve and the city hotel at the city of Lisbon.

```{r}
# Load the dataset

hotel_bookings <- read.csv("hotel_bookings.csv", na.strings="NULL")
View(hotel_bookings)
```


```{r}
# First look to the dataset

str(hotel_bookings) # check the data type of each column

summary(hotel_bookings)

```



```{r}
# Convert character columns into factors

hotel_bookings_new <- as.data.frame(unclass(hotel_bookings),
                       stringsAsFactors = TRUE)
attach(hotel_bookings_new)
```


```{r}
# Convert binary columns "is_canceled" and "is_repeated_guest" into factor

hotel_bookings_new$is_canceled <- as.factor(hotel_bookings_new$is_canceled)

levels(hotel_bookings_new$is_canceled) <- c(0, 1)



hotel_bookings_new$is_repeated_guest <- as.factor(hotel_bookings_new$is_repeated_guest)

levels(hotel_bookings_new$is_repeated_guest) <- c(0, 1)

```

```{r}
# Convert column "arrival_date_year" into factor

hotel_bookings_new$arrival_date_year <- as.factor(hotel_bookings_new$arrival_date_year)

levels(hotel_bookings_new$arrival_date_year) <- c("2015", "2016", "2017")
```

```{r}
# Convert column "children" into numeric

hotel_bookings_new$children <- as.numeric(as.character(hotel_bookings_new$children))
```


```{r}
# Type of each column

str(hotel_bookings_new)

# Summary

summary(hotel_bookings_new)
```



```{r}
# Missing values


# Replacing missing values in children column from the corresponding babies column

n_children <- length(hotel_bookings_new$children)
for (i in 1:n_children) {
  if (is.na(hotel_bookings_new$children[i]))
    hotel_bookings_new$children[i] <- 0
}

# Remove columns "agent" and "company"

index_agent <- which(colnames(hotel_bookings_new)=="agent")
index_company <- which(colnames(hotel_bookings_new)=="company")
hotel_bookings_new = hotel_bookings_new[-c(index_agent, index_company)]

view(hotel_bookings_new)
summary(hotel_bookings_new)
```

This dataset contains information about two different kinds of hotel: City Hotel and Resort Hotel.

```{r}
# Hotel plot: City and Resort

df_hotel <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0, c("hotel")])
df_hotel <- as.data.frame(lapply(df_hotel, function(x) as.data.frame(table(x))))

colnames(df_hotel) <- c("hotel", "frequency")

df_hotel

fig_hotel <- df_hotel %>% plot_ly(labels = ~hotel, values = ~frequency)
fig_hotel <- fig_hotel %>% add_pie(hole = 0.6)
fig_hotel <- fig_hotel %>% layout(title = "Total number of booking for each hotel",  showlegend = T,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig_hotel

```


Both hotels are located in Portugal; this is the reason why most of the guests come from Portugal, as we can see from the map plot below:

```{r}

# Country plot

df_country <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0, c("country")])

df_country <- as.data.frame(lapply(df_country, function(x) as.data.frame(table(x))))

colnames(df_country) <- c('country', 'frequency')

# we use log-scale to see better the different number of guests for each country

fig_country <- plot_ly(df_country, type='choropleth', locations=df_country$country, z=log(df_country$frequency+1), colorscale = 'Viridis' )

fig_country

```

QUANDO

Total bookings in general and for each year

```{r}
# Total bookings for each hotel by month of arrival date

df_months_City <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel', c( "arrival_date_month")])

df_months_Resort <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel', c( "arrival_date_month")])

df_months_City <- as.data.frame(lapply(df_months_City, function(x) as.data.frame(table(x))))

df_months_Resort <- as.data.frame(lapply(df_months_Resort, function(x) as.data.frame(table(x))))

colnames(df_months_City) <- c("arrival_date_month_City", "frequency_City" )

colnames(df_months_Resort) <- c("arrival_date_month_Resort", "frequency_Resort")

df_months_City$arrival_date_month_City <- factor(df_months_City$arrival_date_month_City, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))

df_months_Resort$arrival_date_month_Resort <- factor(df_months_Resort$arrival_date_month_Resort, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_months <- plot_ly()

fig_months <- fig_months %>% add_lines(data=df_months_City, x = ~arrival_date_month_City, y = ~frequency_City, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_months <- fig_months %>% add_lines(data=df_months_Resort, x = ~arrival_date_month_Resort, y = ~frequency_Resort, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_months <- fig_months %>% layout(title = "Total bookings for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total bookings"))

fig_months

```

Year 2015

```{r} 
# Total bookings for each hotel by month (year: 2015)

df_months_City_2015 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2015, c( "arrival_date_month")])

df_months_Resort_2015 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel'& hotel_bookings_new$arrival_date_year==2015, c( "arrival_date_month")])

df_months_City_2015 <- as.data.frame(lapply(df_months_City_2015, function(x) as.data.frame(table(x))))

df_months_Resort_2015 <- as.data.frame(lapply(df_months_Resort_2015, function(x) as.data.frame(table(x))))

colnames(df_months_City_2015) <- c("arrival_date_month_City", "frequency_City" )

colnames(df_months_Resort_2015) <- c("arrival_date_month_Resort", "frequency_Resort")

df_months_City_2015$arrival_date_month_City <- factor(df_months_City_2015$arrival_date_month_City, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))

df_months_Resort_2015$arrival_date_month_Resort <- factor(df_months_Resort_2015$arrival_date_month_Resort, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_months_2015 <- plot_ly()

fig_months_2015 <- fig_months_2015 %>% add_lines(data=df_months_City_2015, x = ~arrival_date_month_City, y = ~frequency_City, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_months_2015 <- fig_months_2015 %>% add_lines(data=df_months_Resort_2015, x = ~arrival_date_month_Resort, y = ~frequency_Resort, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_months_2015 <- fig_months_2015 %>% layout(title = "2015 - Total bookings for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total bookings"))

fig_months_2015

# Note that we do not have bookings from January to June
```


Year 2016

```{r}
# Total bookings for each hotel by month (year: 2016)

df_months_City_2016 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2016, c( "arrival_date_month")])

df_months_Resort_2016 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel'& hotel_bookings_new$arrival_date_year==2016, c( "arrival_date_month")])

df_months_City_2016 <- as.data.frame(lapply(df_months_City_2016, function(x) as.data.frame(table(x))))

df_months_Resort_2016 <- as.data.frame(lapply(df_months_Resort_2016, function(x) as.data.frame(table(x))))

colnames(df_months_City_2016) <- c("arrival_date_month_City", "frequency_City" )

colnames(df_months_Resort_2016) <- c("arrival_date_month_Resort", "frequency_Resort")

df_months_City_2016$arrival_date_month_City <- factor(df_months_City_2016$arrival_date_month_City, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))

df_months_Resort_2016$arrival_date_month_Resort <- factor(df_months_Resort_2016$arrival_date_month_Resort, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_months_2016 <- plot_ly()

fig_months_2016 <- fig_months_2016 %>% add_lines(data=df_months_City_2016, x = ~arrival_date_month_City, y = ~frequency_City, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_months_2016 <- fig_months_2016 %>% add_lines(data=df_months_Resort_2016, x = ~arrival_date_month_Resort, y = ~frequency_Resort, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_months_2016 <- fig_months_2016 %>% layout(title = "2016 - Total bookings for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total bookings"))

fig_months_2016
```


Year 2017

```{r}
# Total bookings for each hotel by month (year: 2017)

df_months_City_2017 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2017, c( "arrival_date_month")])

df_months_Resort_2017 <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel'& hotel_bookings_new$arrival_date_year==2017, c( "arrival_date_month")])

df_months_City_2017 <- as.data.frame(lapply(df_months_City_2017, function(x) as.data.frame(table(x))))

df_months_Resort_2017 <- as.data.frame(lapply(df_months_Resort_2017, function(x) as.data.frame(table(x))))

colnames(df_months_City_2017) <- c("arrival_date_month_City", "frequency_City" )

colnames(df_months_Resort_2017) <- c("arrival_date_month_Resort", "frequency_Resort")

df_months_City_2017$arrival_date_month_City <- factor(df_months_City_2017$arrival_date_month_City, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))

df_months_Resort_2017$arrival_date_month_Resort <- factor(df_months_Resort_2017$arrival_date_month_Resort, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_months_2017 <- plot_ly()

fig_months_2017 <- fig_months_2017 %>% add_lines(data=df_months_City_2017, x = ~arrival_date_month_City, y = ~frequency_City, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_months_2017 <- fig_months_2017 %>% add_lines(data=df_months_Resort_2017, x = ~arrival_date_month_Resort, y = ~frequency_Resort, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_months_2017 <- fig_months_2017 %>% layout(title = "2017 - Total bookings for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total bookings"))

fig_months_2017

# Note that we do not have bookings from September to December
```

Total guests for each year

Year 2015

```{r}
# Total guests for each hotel by month (year: 2015)

df_2015_City <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2015, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2015_City <- df_2015_City %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 

df_2015_Resort <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel' & hotel_bookings_new$arrival_date_year==2015, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2015_Resort <- df_2015_Resort %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 


df_guests_2015_City$arrival_date_month <- factor(df_guests_2015_City$arrival_date_month, levels = c("July", "August", "September", "October", "November", "December"))

df_guests_2015_Resort$arrival_date_month <- factor(df_guests_2015_Resort$arrival_date_month, levels = c("July", "August", "September", "October", "November", "December"))


fig_guests_2015 <- plot_ly()

fig_guests_2015 <- fig_guests_2015 %>% add_lines(data=df_guests_2015_City, x = ~arrival_date_month, y = ~guests, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_guests_2015 <- fig_guests_2015 %>% add_lines(data=df_guests_2015_Resort, x = ~arrival_date_month, y = ~guests, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_guests_2015 <- fig_guests_2015 %>% layout(title = "2015 - Total guests for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total guests"))

fig_guests_2015

```



Year 2016

```{r}
# Total guests for each hotel by month (year: 2016)

df_2016_City <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2016, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2016_City <- df_2016_City %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 

df_2016_Resort <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel' & hotel_bookings_new$arrival_date_year==2016, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2016_Resort <- df_2016_Resort %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 


df_guests_2016_City$arrival_date_month <- factor(df_guests_2016_City$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))

df_guests_2016_Resort$arrival_date_month <- factor(df_guests_2016_Resort$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August", "September", "October", "November", "December"))


fig_guests_2016 <- plot_ly()

fig_guests_2016 <- fig_guests_2016 %>% add_lines(data=df_guests_2016_City, x = ~arrival_date_month, y = ~guests, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_guests_2016 <- fig_guests_2016 %>% add_lines(data=df_guests_2016_Resort, x = ~arrival_date_month, y = ~guests, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_guests_2016 <- fig_guests_2016 %>% layout(title = "2016 - Total guests for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total guests"))

fig_guests_2016
```


Year 2017

```{r}
# Total guests for each hotel by month (year: 2017)

df_2017_City <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='City Hotel' & hotel_bookings_new$arrival_date_year==2017, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2017_City <- df_2017_City %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 

df_2017_Resort <- as.data.frame(hotel_bookings_new[hotel_bookings_new$is_canceled==0 & hotel_bookings_new$hotel=='Resort Hotel' & hotel_bookings_new$arrival_date_year==2017, c( "arrival_date_month", "total_stays", "adults", "children", "babies")])

df_guests_2017_Resort <- df_2017_Resort %>%
  group_by(arrival_date_month) %>%
  summarise(guests = sum(total_stays*(adults+children+babies))) %>%
  ungroup() 


df_guests_2017_City$arrival_date_month <- factor(df_guests_2017_City$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August"))

df_guests_2017_Resort$arrival_date_month <- factor(df_guests_2017_Resort$arrival_date_month, levels = c("January", "February", "March","April", "May", "June", "July", "August"))


fig_guests_2017 <- plot_ly()

fig_guests_2017 <- fig_guests_2017 %>% add_lines(data=df_guests_2017_City, x = ~arrival_date_month, y = ~guests, name = 'City Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)', width = 4))

fig_guests_2017 <- fig_guests_2017 %>% add_lines(data=df_guests_2017_Resort, x = ~arrival_date_month, y = ~guests, name = 'Resort Hotel', type = 'scatter', mode = 'lines', line = list(color = 'rgb(22, 96, 167)', width = 4)) 

fig_guests_2017 <- fig_guests_2017 %>% layout(title = "2016 - Total guests for each hotel by month of arrival date",
         xaxis = list(title = "Months"),
         yaxis = list (title = "Total guests"))

fig_guests_2017
```




#Correlazione

##convert all factors to numeric
tt <- hotel_bookings %>% mutate_if(is.factor, as.numeric)

##calculate correlation coefficient between target variable and the other variables
cor_is_canceled <-  data.frame ( sapply(tt, cor, y=tt$is_canceled))

##raname table
colnames(cor_is_canceled) <- "corr_coeff"

## select only those with corr_coeff < -0.15 | corr_coeff > 0.15
cor_is_canceled <- cor_is_canceled %>%  rownames_to_column(., var = "feature") %>% 
  filter(corr_coeff < -0.15 | corr_coeff > 0.15) %>%
  arrange(desc(corr_coeff))


## subset the tibble with only features that are strongly correlated to the target feature
tt <- tt %>%select(cor_is_canceled$feature)

## print the table
cor_is_canceled  %>% knitr::kable()

cor(tt) %>% ggcorrplot(., hc.order = FALSE, type = "lower", 
                       lab = TRUE)


# Regressione su is_canceled

reg.out <- glm(is_canceled~lead_time+stays_in_week_nights+adults+meal, family='binomial')
summary(reg.out)

# Aggiunta variabile total_stays
hotel_bookings_new=hotel_bookings_new%>%
  mutate(total_stays=(stays_in_week_nights + stays_in_weekend_nights) )

# Regressione completa
model <- glm(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests,  family="binomial")
summary(model)

# Collinearity

VIF(model)

# Selezione del modello
regfit.full=regsubsets(is_canceled~lead_time+total_stays+adults+children+
                       babies+meal+market_segment+distribution_channel+is_repeated_guest+adr+previous_cancellations+
                         previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+
                         required_car_parking_spaces+total_of_special_requests, data = hotel_bookings_new, nvmax = 46)
reg.summary=summary(regfit.full)
names(reg.summary)


par(mfrow=c(2,2))
plot(reg.summary$rss,xlab="n. di predittori",ylab="RSS",type="b",pch=16,cex=.5)
rp=which.min(reg.summary$rss)
points(rp,reg.summary$rss[rp],pch=20,col="red")

plot(reg.summary$adjr2,xlab="n. di predittori",ylab="Adjusted R^2",type="b",pch=16,cex=.5)
rp=which.max(reg.summary$adjr2)
points(rp,reg.summary$adjr2[rp],pch=20,col="red")

plot(reg.summary$cp,xlab="n. di predittori",ylab="Mallow C_p",type="b",pch=16,cex=.5)
rp=which.min(reg.summary$cp)
points(rp,reg.summary$cp[rp],pch=20,col="red")

plot(reg.summary$bic,xlab="n. di predittori",ylab="BIC",type="b",pch=16,cex=.5)
rp=which.min(reg.summary$bic)
points(rp,reg.summary$bic[rp],pch=20,col="red")

par(mfrow=c(1,2))
plot(regfit.full,scale="r2",)
plot(regfit.full,scale="adjr2")
plot(regfit.full,scale="Cp")
plot(regfit.full,scale="bic")

# CV

set.seed(17)
cv.error.10 <- rep(0,10)
for (i in 1:10){
  glm.fit <- glm(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests,  family="binomial",data=hotel_bookings_new)
  cv.error.10[i] <- cv.glm(hotel_bookings_new,glm.fit,K=15)$delta[1]
}
#cv.error.10

par(mfrow=c(1,1))
plot(cv.error.10, type="b",xlab="degree", ylab="K-fold CV error")

# LASSO
library(caret)
set.seed(123)
training.samples <- hotel_bookings_new$is_canceled %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- hotel_bookings_new[training.samples, ]
test.data <- hotel_bookings_new[-training.samples, ]

## Dumy code categorical predictor variables
x <- model.matrix(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests, train.data)[,-1]
## Convert the outcome (class) to a numerical variable
y <- train.data$is_canceled
lambdas <- 10^seq(2, -3, by = -.1)
glmnet(x, y, family = "binomial", alpha = 0, lambda = NULL)

## Find the best lambda using cross-validation
set.seed(123) 
cv.lasso <- cv.glmnet(x, y, alpha = 1, family = "binomial")
## Fit the final model on the training data
model_lasso <- glmnet(x, y, alpha = 1, family = "binomial",
                lambda = cv.lasso$lambda.min)
## Display regression coefficients
coef(model_lasso)
## Make predictions on the test data
x.test <- model.matrix(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests, test.data)[,-1]
probabilities <- model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
## Model accuracy
observed.classes <- test.data$is_canceled
mean(predicted.classes == observed.classes)

cv.lasso <- cv.glmnet(x, y, alpha = 1, family = "binomial")
plot(cv.lasso)
cv.lasso$lambda.min
cv.lasso$lambda.1se
coef(cv.lasso, cv.lasso$lambda.min)
coef(cv.lasso, cv.lasso$lambda.1se)

## Final model with lambda.min
lasso.model <- glmnet(x, y, alpha = 1, family = "binomial",
                      lambda = cv.lasso$lambda.min)
## Make prediction on test data
x.test <- model.matrix(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests, test.data)[,-1]
probabilities <- lasso.model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
## Model accuracy
observed.classes <- test.data$is_canceled
#mean(predicted.classes == observed.classes)

## Fit the model
full.model <- glm(is_canceled~lead_time+reservation_status_date+arrival_date_month+total_stays+adults+children+babies+meal+distribution_channel+is_repeated_guest+adr+previous_cancellations+previous_bookings_not_canceled+booking_changes+deposit_type+days_in_waiting_list+customer_type+required_car_parking_spaces+total_of_special_requests, data = train.data, family = binomial)
## Make predictions
probabilities <- full.model %>% predict(test.data, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
## Model accuracy
observed.classes <- test.data$is_canceled
mean(predicted.classes == observed.classes)

















